---
title: "Homework 3"
author: "Tim Hauser"
output: github_document
---

## Initial setup

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Loading the instacart dataset:

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

In the following I will do an exploration of the  dataset. First, to gain an overview:

```{r}
skimr::skim(instacart)
```

The Instacart Online Grocery Shopping Dataset is an anonymized dataset with over 3 million online grocery orders from more than 200,000 Instacart users from 2017. It consists of `r nrow(instacart)` observations of products orders, and `r ncol(instacart)` different variables: `r names(instacart)`. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Most important variables are:

* `r names(instacart)[1]`: order identifier
* `r names(instacart)[2]`: product identifier
* `r names(instacart)[5]`: customer identifier
* `r names(instacart)[7]`: order sequence number for this user (1=first, n=nth), rages from `r range(pull(instacart, order_number))`
* `r names(instacart)[8]`: day of the week on which the order was placed
* `r names(instacart)[10]`: stands for days since the last order, capped at 30, NA if order_number = 1, ranges from `r range(pull(instacart, days_since_prior_order))`
*`r names(instacart)[10]`: aisle identifier, tells us the location of the item in the warehouse


The following table counts the number of times each aisle is used in the list of orders and orders them by most frequent to least frequent:

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

In total, there are `r instacart %>% select(aisle) %>% distinct %>% count` aisles and the most frequently used are fresh vegetables and fresh fruits.


The following is a plot showing the number of items ordered in each aisle, limited to aisles with more than 10000 items ordered and ordered in ascending order:

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The next table shows the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” and includes number of times each item is ordered in your table:

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

The next table displays the mean hour of the day at which "Pink Lady Apples" and "Coffee Ice Cream" are ordered on each day of the week; formatted in an agenda format (i.e., untidied):

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```


## Problem 2

Load and tidy the dataset (i.e., includes all originally observed variables and values; has useful variable names; includes a weekday vs weekend variable; encodes with reasonable variable classes, reordered variables):
```{r}
accel_df = read.csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity_count",
    names_prefix = "activity_"
  ) %>% 
  rename(day_type_1 = day) %>% 
  mutate(
    day_type_2 = if_else((day_type_1 == "Saturday" | day_type_1 == "Sunday"),"Weekend","Weekday"),
    minute = as.numeric(minute),
    week = as.numeric(week),
    day_id = as.numeric(day_id),
    day_type_1 = factor(day_type_1, level = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
    day_type_2 = factor(day_type_2, level = c("Weekday", "Weekend"))
  ) %>% 
  relocate(week, day_id, day_type_1, day_type_2, minute, activity_count)
```

The following helps to generate an overview of the tidied dataset:

```{r}
skimr::skim(accel_df)
```

The dataset contains five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). Concretely, it consitst of `r nrow(accel_df)` observations of activity counts (one for each minute of each day of a 5 week period), and `r ncol(accel_df)` different variables: `r names(accel_df)` 

* `r names(accel_df)[1]`: week identifier, from `r range(pull(accel_df, week))`, numerical variable
* `r names(accel_df)[2]`: day identifier, from `r range(pull(accel_df, day_id))`, numerical variable
* `r names(accel_df)[3]`: day type 1 from Monday, Sunday, categorical variable
* `r names(accel_df)[4]`: day type 2, Weekday vs. Weekend, categorical variable
* `r names(accel_df)[5]`: minute identifier, from `r range(pull(accel_df, minute))`, numerical variable
* `r names(accel_df)[6]`: activity count (for each minute)

Creating a table with total activity over each day (aggregated across minute):

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  knitr::kable()
```

Plotting above data for better visibility:

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  summarize(total_activity = sum(activity_count), day_type_2 = day_type_2) %>% 
  ggplot(aes(x = day_id, y = total_activity)) + 
  geom_point() + 
  geom_line() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Total activity count per day",
    x = "Day ID",
    y = "Total Activity Counts",
    caption = "Data from the accel dataset"
    )
```

Looking at above graph it's difficult to make out a trend, as total activity counts vary greatly from day to day. The fitted line is decreasing, indicating that total activity count seems to have been going down over the course of the 5 weeks.

In the following I create a similar graph but separate weekdays from weekends:

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  summarize(total_activity = sum(activity_count), day_type_2 = day_type_2) %>% 
  ggplot(aes(x = day_id, y = total_activity, color = day_type_2)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Total activity count per day, separated by weekday vs. weekend",
    x = "Day ID",
    y = "Total Activity Count",
    caption = "Data from the accel dataset"
    ) +
  theme(legend.title=element_blank()) +
  facet_grid(. ~ day_type_2)
```

The above graphs show that while total activity count for weekday generally seemed to have increased over the course of the study, total activity on weekends drastically declined over the same period. Potentially, the test subject forgot to wear the accelerometer on the weekends, since we get two days of complete inactivity.

The following graph is a single-panel plot that shows the 24-hour activity time courses for each day and uses color to indicate day of the week: 

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  ggplot(aes(x = minute, y = activity_count, color = day_type_1)) + 
  geom_line(alpha = 0.5) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Hour of the day",
    y = "Activity Count",
    caption = "Data from the accel dataset"
    ) +
  scale_x_continuous(
    breaks = seq(0,1440,by = 60),
    label = seq(0,24,by = 1)) +
  theme(legend.title=element_blank())
```

First of all, the grah is quite cluttered, making it difficult to deduct a lot from it. What can be said is that there is generally close to zero activity during midnight-5am, very low activity between 5-6am & 10pm-midnight (we can assume this person usually goes to bed at 10pm and wakes up at 6am), 'normal' activity from 6am-7.30pm and a period of high activity between 7.30-10pm. It also becomes evident from the graph that Sundays are quite slow days for this person, with generally low activity throughout the day.   


